---
title: "2020-02-06-HME-peak-annotate-IRanges"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Annotating ChIP-seq peaks!

### Install packages (if not already installed)
```{r}
#BiocManager::install(c("GenomicRanges","rtracklayer", "ChIPseeker", "IRanges", "AnnotationHub"))
```

### Call libraries of packages (install these too if haven't already)
```{r}
library(tidyverse)
library(GenomicRanges)
library(GenomicFeatures)
library(IRanges)
library(openxlsx)
library(rtracklayer)
library(AnnotationHub)
library(dplyr)
```


### Read in GFF files of annotations for Hme
see also https://stackoverflow.com/questions/29253412/finding-intergenic-regions
```{r}
gff <- GenomicFeatures::makeTxDbFromGFF("ChIPQC/GCF_000306765.2_ASM30676v2_genomic.gff", format = "gff")
gff.df <- read_csv("20181113_hfxmed_GCF_000306765.2_ASM30676v2_genomic.gff.key.csv")

#subset whole genome gff into promoter only (i.e. intergenic ranges) and gene only structures
genes.only <- GenomicFeatures::genes(gff)
genes.only.redu <- reduce(genes.only, ignore.strand = T)
ig <- gaps(genes.only.redu)
ig.only <- ig[strand(ig) == "*"] ## This step is important to avoid duplicate entries for each gene.

#pro250 <- GenomicRanges::promoters(genes.only, upstream = 250, downstream = 0)
## Warning: this range of sequences misses large promoters and 3' UTRs (i.e. some IG space is skipped in the overlaps with ChIP-seq peaks)

pro500 <- GenomicRanges::promoters(genes.only, upstream = 500, downstream = 0)
#gene.plus.pro250 <- GenomicRanges::punion(genes.only, pro250)
#gene.plus.pro500 <- GenomicRanges::punion(genes.only, pro500)
```

### Group files containing peak lists that overlap between reps and show differential binding from output of diffbind
In the case of the RosR experiment, this filenames object will only have 1 file because there is only 1 condition with 4 bioreps.
```{r}
filenames <- list.files(path = "peaklists/", pattern = ".gff")
```

### Find overlaps between peak genomic ranges and custom annotation objects (promoters vs genes)
```{r}
for (i in 1:length(filenames)) {
  #import gff file and convert to dataframe
  rosR.pks <- rtracklayer::import(paste("peaklists/", filenames[i], sep = ""), format = "GFF")
  rosR.pks.df <- as.data.frame(rosR.pks)
  
  #add informative meta data
  rosR.pks$peakID <- seq(1,length(rosR.pks))
  rosR.pks$peakwidth <- rosR.pks.df$width
  
  #create hits objects of the overlaps (all peak widths have been standardized to 300bp wide. The must overlap a genomic feature by at least one third (100bps) to be called.) adjusting the overlap requirement changes the stringency of my peak annotation. 
  GenomicRanges::findOverlaps(genes.only, rosR.pks, ignore.strand=T, minoverlap = 100) -> genes
  GenomicRanges::findOverlaps(ig.only, rosR.pks, ignore.strand=T, minoverlap = 100) -> promoters
  GenomicRanges::findOverlaps(pro500, rosR.pks, ignore.strand = T, minoverlap = 100) -> promoters2

  #get IRanges from hits objects and add informative metadata
  genelist <- genes.only[queryHits(genes)] 
  genelist$type <- rep("genic", length(genes))
  genelist$reps <- rosR.pks$name[subjectHits(genes)]
  pintersect(genes.only[queryHits(genes)], rosR.pks[subjectHits(genes)]) -> overlaps
  genelist$overlap <- width(overlaps)
  genelist$peakID <- rosR.pks$peakID[subjectHits(genes)]
  
  prolist <- ig.only[queryHits(promoters)]
  prolist$type <- rep("promoter", length(promoters))
  prolist$reps <- rosR.pks$name[subjectHits(promoters)]
  pintersect(ig.only[queryHits(promoters)], rosR.pks[subjectHits(promoters)]) -> overlaps
  prolist$overlap <- width(overlaps)
  prolist$peakID <- rosR.pks$peakID[subjectHits(promoters)]
  #prolist$gene_id <- NA
  
  prolist2 <- pro500[queryHits(promoters2)]
  prolist2$type <- rep("promoter500", length(promoters2))
  prolist2$reps <- rosR.pks$name[subjectHits(promoters2)]
 pintersect(pro500[queryHits(promoters2)], rosR.pks[subjectHits(promoters2)]) -> overlaps
prolist2$overlap <- width(overlaps)
  prolist2$peakID <- rosR.pks$peakID[subjectHits(promoters2)]
  
  #convert seperate IRanges to Dataframes
  seqs <- seq(1, length(genes))
  as.data.frame(prolist) -> one
  rownames(one) <- NULL
  as.data.frame(genelist, row.names(seqs)) -> two
  rownames(two) <- NULL
  as.data.frame(prolist2) -> three
  rownames(three) <- NULL  
  
  #combine dfs (gene hits and promoter hits)
  #final <- rbind(one, two) %>% arrange(seqnames, start, peakID)
   final <- dplyr::bind_rows(one, two,three) %>% arrange(seqnames, start, peakID)
  #colnames(final)[6] <- "locus_tag"
   colnames(final)[9] <- "locus_tag"

  #merge with gff information (get NCBI annotations and locus names)
  gff.df[gff.df$locus_tag %in% final$locus_tag,] -> tmp



  tmp[c(2,3,4,10)] -> tmp2
  left_join(final, tmp2, by = "locus_tag", na.rm=TRUE) -> final

  
  #reorder
  #final <- final[c(9,1,2,3,7,4,8,5,6,11,10,12)]

  #write to excel file, storing results for each .gff as a different sheet (in the case where multiple peak lists, e.g. between conditions, are used)
    if (i == 1) {
      write.xlsx(final, "peaklists/HME_genelists.xlsx", str_remove_all(filenames[i], ".gff"), row.names = F, asTable = TRUE)
    } else {
        write.xlsx(final, "peaklists/HME_genelists.xlsx", str_remove_all(filenames[i], ".gff"), row.names = F, append = T, asTable = TRUE)}
}
```




```{r}
filenames
```





