---
title: "RosR growth full"
author: "Amy Schmid, Cynthia Darnell"
date: "Oct21,2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}

#fresh start
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## Read in raw data files and tidy

```{r read and tidy}

Raw_untidy <- read.csv("Raw_growth_data_rosR_full.csv", header=TRUE, check.names=FALSE)
#note not all of these data were included in the manuscript due to technical reasons. If you reuse these data please use only the filtered data below (i.e, the "data_filter" object below includes only those were used in the final published manuscript).
Raw <- Raw_untidy %>%
  pivot_longer(cols = `0`:`96`, names_to = "Time", values_to = "OD")

Raw$Time <- as.numeric(Raw$Time)
head(Raw)

```

```{r}

unique(Raw$Experiment)

```

## Subtract blank per experiment (lowest value - 0.02)

```{r}
blank <- Raw %>%
  group_by(interaction(Experiment, Species)) %>%
  summarise(minOD = min(OD, na.rm = TRUE) - 0.002) %>%
  rename(blank_ID = `interaction(Experiment, Species)`)

data <- Raw %>%
  mutate(blank_ID = interaction(Experiment, Species)) %>%
  left_join(blank, by = 'blank_ID') %>%
  mutate(newOD = OD - minOD)


```

## Subset conditions finally used in the ms 
```{r}
data_filter <- filter(data, Experiment == '20150716' | Experiment == '20150721' | Experiment == 'AH_med42'| Experiment == "20150617" |  Experiment == '20150919' |Experiment == '20160316' | Experiment == '20160321' | Experiment == '20141023' | Experiment == '20150607' ) %>%
  filter( Condition == 'standard' | Condition == '0.5 mM H2O2' | Condition == '0.667 mM paraquat' | Condition == "2 mM paraquat" | Condition == "5 mM H2O2" | Condition == "0.333 mM paraquat") %>%
  filter (! Strain == "blank")



unique(data_filter$Strain)
unique(data_filter$Condition)
unique(data_filter$Species)
```


### Format for bsd

``` {r echo = FALSE}
#pivot to wide format
reuntidy <- data_filter %>%
  dplyr::select(-c(OD, minOD)) %>%
  pivot_wider(names_from = Time, values_from = newOD)

output<-dplyr::select (reuntidy, -c(ID, Speed,blank_ID )) 
#write output file for supplementary table 3 of paper
write.csv(output, "Table_S3_growth.csv")

bsd_data <- reuntidy %>%
  dplyr::select(-c(Experiment, Well, Species, Strain, Biological_replicate, 
            Technical_replicate, Speed, Media, Condition, blank_ID))

bsd_data_colnames <- bsd_data$name
bsd_data_rownames <- bsd_data$ID
bsd_data_t <- as.data.frame(t(bsd_data[, -1]))
colnames(bsd_data_t) <- bsd_data_colnames %>%
  as.numeric()
colnames(bsd_data_t) <- bsd_data_rownames

bsd_data_t <- tibble::rownames_to_column(bsd_data_t, "Time")


# Trim to 48 hours for bsd

bsd_48 <- bsd_data_t[1:97,]

#remove first 3 hours of each time course to avoid 0 subraction errors in mumax 

bsd_48_remove3 <- subset (bsd_48, Time > 3.0)
write.csv (bsd_48_remove3, 'bsd_input_final_3plus.csv', row.names = FALSE)

#write.csv(bsd_48, 'bsd_input_final_AKS.csv', row.names = FALSE)

```

## Source bsd and extract xn.auc
### Rmd spits out errors that console doesn't, so you may want to copy/paste lines
```{r}
#make sure bsdAnalysisFunc.R is in working directory

source('bsdAnalysisFunc.R')

bsd <- bsdProcess('bsd_input_final_3plus.csv', convert.time = FALSE)
result_AUC <- apply(bsd$xn, 2, function(y) {simp(y, x=bsd$at)})
mumax <- getMuMax(bsd$mu)
#result <- cbind(xn.auc, mumax)
write.csv(result_AUC, 'AUC_bsd_result_AKS.csv', row.names = TRUE)
write.csv(mumax, "mumax_bsd_result_AKS.csv", row.names = TRUE)

```

## Merge AUC back to data frame to provide metadata for each curve unique ID

```{r}

result_AUC_frame <- as.data.frame(result_AUC)
result_AUC_frame <- tibble::rownames_to_column(result_AUC_frame, "ID")

data_AUC <- merge(data_filter, result_AUC_frame, by = 'ID', all.y = TRUE)

#compress the data frame such that there is only one AUC value per ID growth curve (rather than a repeated AUC per each time point)
auc.unique  <- data_AUC[!duplicated(data_AUC$result_AUC ), ]
auc.unique <- select (auc.unique, -Time)
write.csv(auc.unique,"AUC_all_AKS.csv" )

```

## Merge mumax back to data frame to provide metadata for each curve unique ID

```{r}

result_mu_frame <- as.data.frame(mumax)
result_mu_frame <- tibble::rownames_to_column(as.data.frame(mumax), "ID")

data_mu <- merge(data_filter, result_mu_frame, by = 'ID', all.y = TRUE)

#compress the data frame such that there is only one mu value per ID growth curve (rather than a repeated AUC per each time point)
mu.unique  <- data_mu[!duplicated(data_mu$mumax ), ]
mu.unique <- select (mu.unique, -Time)
write.csv(mu.unique,"mu_all_AKS.csv" )

```

## calculate stats
```{r}
#remove unnecesary columns
auc.final <- select (auc.unique, -c(ID, Experiment, Well, Speed, Media, OD, blank_ID, minOD, newOD))
mu.final <- select (mu.unique, -c(ID, Experiment, Well, Speed, Media, OD, blank_ID, minOD, newOD))

#auc stats
stats.auc <- plyr::ddply (auc.final, c("Strain", "Condition", "Species"), summarise, avg.auc = mean (result_AUC), sd = sd (result_AUC), number.observations = length (result_AUC), sem = (sd(result_AUC))/sqrt(number.observations))

stats.mu <- plyr::ddply(mu.final, c("Strain", "Condition", "Species"), summarise, avg.mu = mean (mumax), sd = sd (mumax), number.observations = length (mumax), sem = (sd(mumax))/sqrt(number.observations))
#mu stats

#write output files
write.csv (stats.auc, file = "stats_AUC_all_AKS.csv")
write.csv (stats.mu, file = "stats_mu_all_AKS.csv")
```

### Plot AUCs by species

```{r}

hv_AUC <- filter(stats.auc, Species == 'Haloferax volcanii' ) %>%
  filter (!Condition == "0.333 mM paraquat")

#pdf("Hv_AUC_allconds_AKS.pdf", width = 6, height = 3)
ggplot(hv_AUC, aes(x = Strain, y = avg.auc, fill = Strain)) +
  #geom_point(aes(color = Strain, group = ID)) +
  geom_bar( stat = "identity", position_dodge = 0.9, alpha = 0.5) +
  theme_light() +
   scale_fill_manual(values = c("black", "salmon")) +
    geom_errorbar(aes(ymin = avg.auc - sem, ymax = avg.auc + sem),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
  labs(y = "AUC", title = "Hfx. volcanii") +
  facet_wrap(~Condition) +
  #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
   #     strip.background = element_rect(fill = 'black'), 
    #    legend.position = 'bottom') +
  #scale_color_viridis_d('Strain', begin = 0, end = 0.7) +
  #scale_fill_viridis_d('Strain', begin = 0, end = 0.7) +
    
  NULL
#dev.off()


#ggsave("Hv_AUC_all.pdf", width = 8.5, height = 11)

```

```{r}

hm_AUC <- filter(stats.auc, Species == 'Haloferax mediterranei' ) 
  
#pdf("Hm_AUC_allconds_AKS.pdf", width = 6, height = 3)
ggplot(hm_AUC, aes(x = Strain, y = avg.auc, fill = Strain)) +
  #geom_point(aes(color = Strain, group = ID)) +
  geom_bar( stat = "identity", alpha = 0.5) +
  theme_light() +
   scale_fill_manual(values = c("black", "salmon")) +
    geom_errorbar(aes(ymin = avg.auc - sem, ymax = avg.auc + sem),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
  labs(y = "AUC", title = "Hfx. mediterranei") +
  facet_wrap(~Condition) +
  #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
   #     strip.background = element_rect(fill = 'black'), 
    #    legend.position = 'bottom') +
  #scale_color_viridis_d('Strain', begin = 0, end = 0.7) +
  #scale_fill_viridis_d('Strain', begin = 0, end = 0.7) +
    
  NULL
#dev.off()




```

```{r}
#Hbt salinarum 
hs_AUC <- filter(stats.auc, Species == 'Halobacterium salinarum' ) 
  
#pdf("Hs_AUC_allconds_AKS.pdf", width = 6, height = 3)
ggplot(hs_AUC, aes(x = Strain, y = avg.auc, fill = Strain)) +
  #geom_point(aes(color = Strain, group = ID)) +
  geom_bar( stat = "identity", alpha = 0.5) +
  theme_light() +
   scale_fill_manual(values = c("salmon", "black")) +
    geom_errorbar(aes(ymin = avg.auc - sem, ymax = avg.auc + sem),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
  labs(y = "AUC", title = "Hbt. salinarum") +
  facet_wrap(~Condition) +
  #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
   #     strip.background = element_rect(fill = 'black'), 
    #    legend.position = 'bottom') +
  #scale_color_viridis_d('Strain', begin = 0, end = 0.7) +
  #scale_fill_viridis_d('Strain', begin = 0, end = 0.7) +
    
  NULL
#dev.off()




```


#plot to see error bars for rosR vs pyrE2 in Hv
```{r}
 pdf("FigS1_HvAUC_0.5mMH2O2_barplot.pdf")
ggplot(subset(as.data.frame (auc.stats.unique), Condition == "0.5 mM H2O2"), aes(x = Strain, y = avg.auc, fill = Condition)) +
  xlab("") +
  ylab("AUC") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg.auc - sd, ymax = avg.auc + sd),
    width = .1,
    size = 1, position = position_dodge(0.9)
  ) +
  
  NULL
dev.off()
```


```{r}
#ttest
H2O2 <- subset(auc.unique, Condition == "0.5 mM H2O2" )
H2O2.pyr <- subset (H2O2, Strain == "pyrE2")
H2O2.ros <- subset (H2O2, Strain == "rosR")
t.test (H2O2.pyr$result_AUC, H2O2.ros$result_AUC, paired = FALSE)
#rosR growth impairment relative to pyrE2 parent under 0.5mM H2O2 is not significant.
```